---
title: "Analogous PT analysis"
author: "Kathryn Denning"
date: "7/6/2020"
output: 
  html_document:
    code_folding: "hide"
    toc: true
    toc_float: true
    dev: png
  pdf_document:
    dev: cairo_pdf
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---

```{r set-up, include = FALSE}
#loading packages
library(foreign)
library(reshape2)
library(tidyverse)
library(dplyr)
library(car)
library(haven)
library(psych)
library(lmerTest)
library(sjPlot)
library(reshape2)
library(data.table)
library(emmeans)
library(effects)
library(Cairo)
library(tinytex)
library(rio)

#making it round to three decimal places
options(scipen = 999)
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

knitr::opts_knit$set(dev.args = list(type = "cairo"))

trace(grDevices:::png, quote({
  if (missing(type) && missing(antialias)) {
    type <- "cairo-png"
    antialias <- "subpixel"
  }
}), print = FALSE)

data <- import("analog_pt_combined_ptfiltered.xlsx")
```

```{r data import, include = FALSE}
# Reverse coding
data[,c("bfi_self_quiet", "bfi_self_rude", 
                 "bfi_self_disorganized", "bfi_self_tasks",
                 "bfi_self_stable", "bfi_self_abstract",
                 "bfi_self_risks", "bfi_self_uses",
                 "bfi_targ_quiet", "bfi_targ_rude",
                 "bfi_targ_disorganized",
                 "bfi_targ_tasks", "bfi_targ_stable",
                 "bfi_targ_abstract", "bfi_targ_risks",
                 "bfi_targ_uses")] <- lapply(data[,c("bfi_self_quiet", "bfi_self_rude", 
                 "bfi_self_disorganized", "bfi_self_tasks",
                 "bfi_self_stable", "bfi_self_abstract",
                 "bfi_self_risks", "bfi_self_uses",
                 "bfi_targ_quiet", "bfi_targ_rude",
                 "bfi_targ_disorganized",
                 "bfi_targ_tasks", "bfi_targ_stable",
                 "bfi_targ_abstract", "bfi_targ_risks",
                 "bfi_targ_uses")], function(x) 
                 recode(x,"'1'=5;
                           '2'=4;
                           '4'=2;
                           '5'=1"))

# Cleaning and putting data in long-format
data_clean <- data %>% 
  mutate(ID = row_number()) %>% 
  # Dropping demographic open-ended questions
  select(-c(gender_write, race_write)) %>% 
  # Putting responses for personality items into long format for participant and target
  pivot_longer(c(bfi_self_quiet:bfi_self_rules, 
           bfi_targ_quiet:bfi_targ_rules),
           names_sep = "_",
           names_to = c("drop1", "bfi_type", "bfi2xsh_qtype")) %>% 
  pivot_wider(names_from = bfi_type, values_from = value) %>% 
  # getting a condition variable
  pivot_longer(c(control_only, imagine_self_only, analog_forcontrol, analog_forself),
               names_to = "condition") %>% 
  mutate(condition = as.factor(condition)) %>% 
  select(-c(control_afteranalog, imagine_self_afteranalog, drop1)) %>% 
  na.omit() %>% 
  select(-value) %>% 
  unique() %>% 
  # Making a variable for perceived polarization
  mutate(perceived_polar = abs(polar_liberal-polar_cons)) %>% 
  # Mean centering continuous predictors
  mutate(self_c = self - mean(self, na.rm = TRUE),
         ident_c = ident - mean(ident, na.rm = TRUE),
         perceived_polar_c = perceived_polar - 
           mean(perceived_polar, na.rm = TRUE)) %>% 
# Removing people who responded "neither" because only want in and out-group for this analysis, this was an error in data collection to have a third option
  filter(cand_pref != 3) %>% 
  #effects coding candidate; Clinton = .5, Trump = -.5
  mutate(cand_pref = recode(cand_pref, "'1' = '.5'; '2' = '-.5'")) %>% 
  #making demographics factors or numeric
  mutate(gender = as.factor(gender),
         race = as.factor(race),
         age = as.numeric(age)) %>% 
  #labeling gender and race
  mutate(gender = recode(gender, "'1' = 'Female';'2' = 'Male'; c('3','4','5')  = 'Other'"),
         race = recode(race, "'1' = 'American Indian or Alaska Native'; '8' = 'Asian'; '9' = 'Black';
                       '10' = 'Latinx'; '12' = 'Middle Eastern'; '14' = 'Hawaiian/Pacific Islander'; '16' = 'White'; c('17', '18') = 'Other'"))
```

# Demographics

## Participant number

```{r participants number}
data_clean %>% 
  select(ID) %>% 
  unique() %>% 
  nrow()
```

## Participants per condition

```{r participants per condition}
data_clean %>% 
  select(ID, condition) %>% 
  unique() %>% 
  group_by(condition) %>% 
  count()
```

## Gender

```{r gender}
data_clean %>% 
  select(ID, gender) %>% 
  unique() %>% 
  group_by(gender) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(n_total = sum(n),
         percent = (n/n_total)*100)
```

## Race

```{r race}
data_clean %>% 
  select(ID, race) %>% 
  unique() %>% 
  group_by(race) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(n_total = sum(n),
         percent = (n/n_total)*100)
```

## Age

```{r age}
data_clean %>% 
  select(ID, age) %>% 
  na.omit() %>% 
  summarize(mean = mean(age),
            sd = sd(age))
```

## Candidate preference percentage

```{r}
data_clean %>% 
  mutate(cand_pref = as.factor(cand_pref)) %>% 
  mutate(cand_pref = recode(cand_pref, "'0.5' = 'Clinton'; '-0.5' = 'Trump'")) %>% 
  select(ID, cand_pref) %>% 
  na.omit() %>% 
  unique() %>% 
  group_by(cand_pref) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(n_total = sum(n),
         percent = (n/n_total)*100)
```

# Overall descriptives

```{r proj descriptives}
describe(data_clean)
```

## Histogram of in-group identification

```{r hist identification}
hist_ident <- data_clean %>% 
  select(ID, ident) %>% 
  unique()
hist(hist_ident$ident)
```

## Histogram of perceived polarization

```{r perceived polarization distribution}
hist_polar <- data_clean %>% 
  select(ID, perceived_polar) %>% 
  unique()
hist(hist_polar$perceived_polar)
```

# Projection Model

## Descriptives: Condition predicting projection

### Mean BFI scores for self by condition

```{r BFI self}
data_clean %>% 
  select(ID, condition, self) %>% 
  na.omit() %>% 
  group_by(condition) %>% 
  summarize(mean = mean(self),
            sd = sd(self))
```

### Mean BFI scores for target by condition

```{r BFI targ}
data_clean %>% 
  select(ID, condition, targ) %>% 
  na.omit() %>% 
  group_by(condition) %>% 
  summarize(mean = mean(targ),
            sd = sd(targ))
```

## Analysis

```{r projection analysis}
#Adding contrast codes for condition
analog_v_non <- c(-1, -1, 1, 1) 
analogc_v_analogs <- c(-1, 1, 0, 0)  
con_v_self <- c(0, 0, 1, -1) 
ConCodes <- cbind(analog_v_non, analogc_v_analogs, con_v_self)
contrasts(data_clean$condition) <- ConCodes
contrasts(data_clean$condition)

proj_model <- lmer(targ ~ self_c*ident_c*condition*cand_pref*perceived_polar_c + (self_c|ID), 
                            data = data_clean)

tab_model(proj_model,
          title = "Projection model")
```

## Plots

```{r main effect identification}
#Getting stats for plot
describe(data_clean$ident_c)

proj_ident <- as_tibble(effect(term = "self_c:ident_c", 
                                 xlevels = list(ident_c = c(-0.95, 0, 0.95)),
                                 mod = proj_model), na.rm = TRUE)

proj_ident$Identification <-as.factor(proj_ident$ident_c)

# Color plot for conference presentation
ggplot(proj_ident, aes(self_c, fit, color = Identification)) +
  geom_smooth(method = "lm", 
              se = FALSE,
              size = 1.25, 
              alpha = .3,
              aes(color = Identification)) +
  scale_color_manual(breaks = c(-0.95, 0, 0.95), 
                     labels = c ("Weak \n(1 SD below)", 
                                  "Average", 
                                  "High \n(1 SD above)"),
                     values = c('#404788FF',
                               '#55C667FF',
                               '#FDE725FF')) +
  theme_minimal(base_size = 13) +
  labs(title = "In-group identification predicting projection",
       x = "Personality responses for self",
       y = "Personality responses for target") +
  ylim(2,4)
```

No one is projecting in this sample... As in-group identification increases, people go from not projecting to their out-group to counter projecting.

```{r main effect polarization}
#Getting stats for plot
describe(data_clean$perceived_polar_c)

proj_polar <- as_tibble(effect(term = "self_c:perceived_polar_c", 
                                 xlevels = list(perceived_polar_c = c(-1.8, 0, 1.8)),
                                 mod = proj_model), na.rm = TRUE)

proj_polar$Polarization <-as.factor(proj_polar$perceived_polar_c)

# Color plot for conference presentation
ggplot(proj_polar, aes(self_c, fit, color = Polarization)) +
  geom_smooth(method = "lm", 
              se = FALSE,
              size = 1.25, 
              alpha = .3,
              aes(color = Polarization)) +
  scale_color_manual(breaks = c(-1.8, 0, 1.8), 
                     labels = c ("Low \n(1 SD below)", 
                                  "Average", 
                                  "High \n(1 SD above)"),
                     values = c('#404788FF',
                               '#55C667FF',
                               '#FDE725FF')) +
  theme_minimal(base_size = 13) +
  labs(title = "Perceived polarization predicting projection",
       x = "Personality responses for self",
       y = "Personality responses for target") +
  ylim(2,4)
```

People who don't perceive polarization do not project. As perceived polarization increases, counter projection increases.

```{r identification by candidate pref}
#Getting stats for plot

proj_identxcand <- as_tibble(effect(term = "self_c:ident_c:cand_pref", 
                                 xlevels = list(ident_c = c(-0.95, 0, 0.95),
                                                cand_pref = c("0.5", "-0.5")),
                                 mod = proj_model), na.rm = TRUE)

proj_identxcand$Identification <-as.factor(proj_identxcand$ident_c)
proj_identxcand$Candidate <-as.factor(proj_identxcand$cand_pref)

ident_label <- c("-0.95" = "Low identification \n (1 SD below)", 
                 "0" = "Average identification",
                 "0.95" = "High identificatiion \n(1 SD above)")

# Color plot for conference presentation
ggplot(proj_identxcand, aes(self_c, fit, color = Candidate)) +
  geom_smooth(method = "lm", 
              se = FALSE,
              size = 1.25, 
              alpha = .3,
              aes(color = Candidate)) +
  facet_wrap(~Identification, labeller = 
               labeller(Identification = ident_label)) +
  scale_color_manual(breaks = c(-0.5, 0.5), 
                     labels = c ("Clinton", 
                                  "Trump"),
                     values = c('#404788FF',
                               '#55C667FF')) +
  theme_minimal(base_size = 13) +
  labs(title = "In-group identification and candidate predicting projection",
       x = "Personality responses for self",
       y = "Personality responses for target") +
  ylim(2,4)
```

Low identifying Clinton supporters project normally. However, as in-group identification increases, projection decreases and ultimately becomes counter projection. Trump supporters counter project to a similar degree regardless of identification. 


```{r candidate pref by pt}
#Getting stats for plot
proj_candxcond <- as_tibble(effect(term = "self_c:condition:cand_pref", 
                                 xlevels = list(cand_pref = c("0.5", "-0.5")),
                                 mod = proj_model), na.rm = TRUE)

proj_candxcond$Candidate <-as.factor(proj_candxcond$cand_pref)

cand_label <- c("-0.5" = "Trump", 
                 "0.5" = "Clinton")


ggplot(proj_candxcond, aes(self_c, fit, color = condition)) +
  geom_smooth(method = "lm", 
              se = FALSE,
              size = 1.25, 
              alpha = .3,
              aes(color = condition)) +
  facet_wrap(~Candidate, labeller = 
               labeller(Candidate = cand_label)) +
  scale_color_manual("Condition",
                     labels = c("Analog + Control",
                                 "Analog + Imagine Self",
                                 "Control Only",
                                 "Imagine Self Only"),
                     values = c('#404788FF',
                               '#238A8DFF',
                               '#55C667FF',
                               '#FDE725FF')) +
  theme_minimal(base_size = 13) +
  labs(title = "PT condition and candidate predicting projection",
       subtitle = "Comparison between analogous versus non-analogous is significant",
       x = "Personality responses for self",
       y = "Personality responses for target") +
  ylim(2,4)
```

It appears that adding analogous perspective taking to imagine-self instructions leads Trump supporters to counter project more than in the imagine self only condition as well as more than Clinton supporters in the analogous + imagine self condition. Interestingly, Trump supporters positively projected in the control only condition, while that is the condition that Clinton supporters counter projected the least in. Adding analogous instructions to control instructions for Trump supporters then lead to counter projection similiar to the imagine self only condition. Both the ananlogous conditions lead to similiar counter projection for Clinton supporters, while Clinton supporters counter projected the most in the imagine self only condition.

# Perspective Taking Model

## Descriptives: Condition predicting perspective taking

```{r pt by condition}
pt_cond <- data_clean %>% 
  select(ID, condition, pt_mancheck) %>% 
  na.omit() %>% 
  group_by(condition) %>% 
  summarize(mean = mean(pt_mancheck),
            sd = sd(pt_mancheck),
            se = sd(pt_mancheck)/sqrt(length(pt_mancheck))) %>% 
  mutate(condition = recode(condition,
                            "'analog_forcontrol' = 'Analog + Control';
                            'analog_forself' = 'Analog + Imagine Self';
                            'control_only' = 'Control Only';
                            'imagine_self_only' = 'Imagine Self Only'"))
pt_cond
```

## Analysis

```{r pt model}
pt_data <- data_clean %>% select(c(ID, pt_mancheck, ident_c, condition, cand_pref, perceived_polar_c)) %>% 
  unique() %>% 
  na.omit()

pt_model <- lm(pt_mancheck ~ ident_c*condition*cand_pref*perceived_polar_c, 
                            data = pt_data)

tab_model(pt_model,
          title = "Perspective taking model")
```

## Plots

```{r condition main predicting pt}
ggplot(pt_cond, aes(condition, mean, fill = condition)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(title = "Self-reported perspective taking predicted by perspective taking instructions",
       x = "Instructions",
       y = "Self-reported perspective taking") +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = .2,
                position = position_dodge(.9)) +
  scale_fill_viridis_d()
```

Results indicate that people in the imagine-self conditions report taking the perspective of their out-group target less than those who are in the control group, regardless of analogous instructions. Possibly a back firing effect due to threat? The more you are told to consider the perspective of an out-group target, the less you want to.

```{r ident x cand_pref predicting pt}
ptxidentxcand <- data_clean %>% 
  select(ident_c, pt_mancheck, cand_pref, ID) %>% 
  mutate(cand_pref = recode(cand_pref, "'0.5' = 'Clinton'; '-0.5' = 'Trump'"))

ggplot(ptxidentxcand, aes(ident_c, pt_mancheck, group = cand_pref, color = cand_pref)) +
  geom_smooth(method = "lm",
              se = FALSE) +
   scale_color_manual("Candidate",
                     values = c('#404788FF',
                               '#55C667FF')) +
  theme_minimal(base_size = 13) +
  labs(title = "In-group identification and candidate predicting self-reported perspective taking",
       x = "In-group identification",
       y = "Self-reported perspective taking")
```

As in-group identification increases, Trump supporters report taking the perspective of the out-group target more, while Clinton supporters report doing so less.
